---
pagetitle: "COVID19"
output:
  html_document:
    includes:
       in_header: header.html
---

  
  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# MIT Licence
# Copyright 2020 Martin Spendiff
# 
# Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the "Software"), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:
# 
# The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.
# 
# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
#SETUP


rm(list=ls(all=TRUE)) 
library(ggplot2)
library(knitr)
library(dplyr)
library(lubridate)
library(stringr)
library(plotly)
library(DT)
library(jsonlite)
library(reshape2)
library(scales)

# FUNCTIONS

toperson <- function(personnum, colourstr){
unlucky<-sample(1:personnum, 1, replace=T)
people<-paste(strrep(paste("<i class=\"fas fa-user\" style=\"color:",colourstr,"\"></i> ", sep=""), unlucky-1),
      strrep("<i class=\"fas fa-user\"  style=\"color:red\"></i> ",1),
      strrep(paste("<i class=\"fas fa-user\" style=\"color:",colourstr,"\"></i> ", sep=""), personnum-unlucky),
      strrep(" ",2+personnum), sep="")
return(people)
}

###################################### End of Functions

# Parameters

topcount<-10
confirmed.cutoff<-30
coi<-sort(c("China", "Italy", "United Kingdom","Switzerland", "Germany","Norway","South Africa","US","France","Japan", "New Zealand", "Spain"))
minimum.pop=5e6
urlcases<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master"
casespercapita.origin<-1e-6
casespercapita.cutoff<-5000000


###################################### End of Parameters

# Get data

# Time series data (Source reliability seems to be deteriorating. Consider changing if this is not rectified.)
 
death.wide <- read.csv(paste(urlcases,"/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv",sep=""))

recovered.wide <- read.csv(paste(urlcases,"/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv",sep=""))

confirmed.wide <- read.csv(paste(urlcases,"/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv", sep=""))

# csv from high frequency worldometer scrape. 

worldometer<- read.csv("/Users/martin/Desktop/therepo/nCoV19/R.Script/output_file.csv")

# Get population data from the World Bank API
url <- 'http://api.worldbank.org/v2/country/all/indicator/SP.POP.TOTL?date=2018&format=json&per_page=300'

# read url and convert to data.frame
population.doc <- fromJSON(txt=url)

# Continent -> Country dataset downloaded from Kaggle
country2continent <- read.csv(
  "https://spendiff.com/countryContinent.csv")

####################################################### End of getting data


# Tidy and rearrange data

# Turn date column titles into something I can work with Currently XM.DD.YY
# Convert to DD.MM.YY

total.cases.worldometer<-as.numeric(sub(",", "",as.character(worldometer[worldometer$Country.Other=="Total:",]$TotalCases)), fixed=T)
total.deaths.worldometer<-as.numeric(sub(",", "",as.character(worldometer[worldometer$Country.Other=="Total:",]$TotalDeaths)), fixed=T)
total.recovered.worldometer<-as.numeric(sub(",", "",as.character(worldometer[worldometer$Country.Other=="Total:",]$TotalRecovered)), fixed=T)

names(death.wide)[4:length(names(death.wide))]<-str_replace(names(death.wide[4:length(names(death.wide))]), "X(\\d\\d?)+.(\\d\\d?).(\\d\\d)", "\\2.\\1.\\3")
names(recovered.wide)[4:length(names(recovered.wide))]<-str_replace(names(recovered.wide[4:length(names(recovered.wide))]), "X(\\d\\d?)+.(\\d\\d?).(\\d\\d)", "\\2.\\1.\\3")
names(confirmed.wide)[4:length(names(confirmed.wide))]<-str_replace(names(confirmed.wide[4:length(names(confirmed.wide))]), "X(\\d\\d?)+.(\\d\\d?).(\\d\\d)", "\\2.\\1.\\3")

# Now melt them

death.long <- melt(death.wide, id=c("Province.State","Country.Region","Lat","Long" ))
names(death.long)[5]<-"Date"
names(death.long)[6]<-"number.death"
recovered.long <- melt(recovered.wide, id=c("Province.State","Country.Region","Lat","Long" ))
names(recovered.long)[5]<-"Date"
names(recovered.long)[6]<-"number.recovered"
confirmed.long <- melt(confirmed.wide, id=c("Province.State","Country.Region","Lat","Long" ))
names(confirmed.long)[5]<-"Date"
names(confirmed.long)[6]<-"number.confirmed"

population.data<-as.data.frame(population.doc[[2]])
names(population.data)[3]<-"code_3"
population.data<-merge(population.data,country2continent, by=c("code_3"))

population<-as.data.frame(cbind("Country.Region"=population.data$country.x$value,"population"=population.data$value,"year"=population.data$date,"continent"=as.character(population.data$continent)),stringsAsFactors = FALSE)


################################################## Repair documented data errors on the github dataset
taiwan<-c("Taiwan*","32780000","2018","Asia")
population<-rbind (population, taiwan) 

########## March 12th issues REMOVE AFTER FIXED AT SOURCE (REMOVED!)
#
#Italy
#Confirmed: 15113
#Deaths: 1016
#Recovered: 1258
death.long$number.death[death.long$Country.Region==("Italy") & death.long$Date==("12.3.20")] <-1016
confirmed.long$number.confirmed[confirmed.long$Country.Region==("Italy") & confirmed.long$Date==("12.3.20")] <-15113
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

#UK
confirmed.long$number.confirmed[confirmed.long$Country.Region==("United Kingdom") & confirmed.long$Province.State==("United Kingdom")  & confirmed.long$Date==("12.3.20")] <-561
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

#Switzerland
death.long$number.death[death.long$Country.Region==("Switzerland") & death.long$Date==("12.3.20")] <-7
confirmed.long$number.confirmed[confirmed.long$Country.Region==("Switzerland") & confirmed.long$Date==("12.3.20")] <-868
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

#Norway
death.long$number.death[death.long$Country.Region==("Norway") & death.long$Date==("12.3.20")] <-1
#confirmed.long$number.confirmed[confirmed.long$Country.Region==("Norway") & confirmed.long$Date==("12.3.20")] <-817

#France
death.long$number.death[death.long$Country.Region==("France") & death.long$Date==("12.3.20")][1] <-61
confirmed.long$number.confirmed[confirmed.long$Country.Region==("France") & confirmed.long$Date==("12.3.20")][1] <-2876
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

#Germany
death.long$number.death[death.long$Country.Region==("Germany") & death.long$Date==("12.3.20")][1] <-6
confirmed.long$number.confirmed[confirmed.long$Country.Region==("Germany") & confirmed.long$Date==("12.3.20")][1] <-2750
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

#Iran
death.long$number.death[death.long$Country.Region==("Iran") & death.long$Date==("12.3.20")][1] <-429
confirmed.long$number.confirmed[confirmed.long$Country.Region==("Iran") & confirmed.long$Date==("12.3.20")][1] <-10075
#recovered.long$number.recovered[recovered.long$Country.Region==("Italy") & recovered.long$Date==("12.3.20")]<-1258

confirmed.long$number.confirmed[confirmed.long$Country.Region==("Japan") & confirmed.long$Date==("6.2.20")][1]<-23

#############################


# Get counts befor the merges so that disappearing countries will not deplete totals

df <-merge(death.long, confirmed.long, by=c("Lat","Long","Date","Country.Region","Province.State"))
df <-merge(df, recovered.long, by=c("Lat","Long","Date","Country.Region","Province.State"))
df$Date<-dmy(df$Date)

################ Where multiple regions for the same country are listed on a date, sum and summarise 

df<- df %>%
   group_by(Country.Region, Date) %>%
   summarise(number.confirmed=sum(number.confirmed),number.death=sum(number.death),number.recovered=sum(number.recovered))

###################################################################### End of region merge

growth<-df

growth<-growth %>%
       group_by(Date)%>%
       summarise("confirmed"=sum(number.confirmed),"death"=sum(number.death),"recovered" = sum(number.recovered)) %>%
       mutate("newconfirmed"=confirmed-lag(confirmed),"newdeath"=death-lag(death),"newrecovered"=recovered-lag(recovered)) %>%
       ungroup()

growth.by.country<-df %>%
       group_by(Date,Country.Region) %>%
       summarise("confirmed"=sum(number.confirmed),"death"=sum(number.death),"recovered" = sum(number.recovered)) %>%
       ungroup()

growth.by.country<-growth.by.country %>%
       group_by(Country.Region) %>%
       mutate("newconfirmed"=confirmed-lag(confirmed),"newdeath"=death-lag(death),"newrecovered"=recovered-lag(recovered),"active"=confirmed-death-recovered)%>%
       ungroup()

###################################################################### End of get counts


# Fix of Some countrynames for alignment between data sets

population$Country.Region[population$Country.Region=="Korea, Rep."]<-"Korea, South"
population$Country.Region[population$Country.Region=="United States"]<-"US"
#population$Country.Region[population$Country.Region=="United Kingdom"]<-"UK"
population$Country.Region[population$Country.Region=="Iran, Islamic Rep."]<-"Iran"
#population$Country.Region[population$Country.Region=="Hong Kong SAR, China"]<-"Hong Kong SAR"
population$Country.Region[population$Country.Region=="Viet Nam"]<-"Vietnam"
population$Country.Region[population$Country.Region=="Slovak Republic"]<-"Slovakia"
population$Country.Region[population$Country.Region=="Russian Federation"]<-"Russia"
population$Country.Region[population$Country.Region=="Egypt, Arab Rep."]<-"Egypt"
population$Country.Region[population$Country.Region=="Czech Republic"]<-"Czechia"
# unique(df$Country.Region[!((df$Country.Region) %in% population$Country.Region)]) # Countries and data get lost in the population merge. Fix this


# Merge death recovered and confirmed
df.pre.pop.merge<-df
df<- merge(df, population, by=c("Country.Region"))




df$Country.Region<-as.character(df$Country.Region)
df$population<-as.integer(df$population)

df$number.death.per.capita<-df$number.death/df$population
df$number.recovered.per.capita<-df$number.recovered/df$population
df$number.confirmed.per.capita<-df$number.confirmed/df$population

# Add a days of data column

mindate<-min(df$Date)
maxdate<-max(df$Date)
df<-df %>% 
    group_by(Date) %>%
    mutate("days.of.data"=difftime(Date, mindate,unit="days")) %>%
    ungroup()

df$continent<-as.factor(df$continent)

# For normalised plots

onsetday<-subset(df,population>casespercapita.cutoff) %>% 
     group_by(Country.Region) %>%
     mutate("day.of.onset"= if_else(number.confirmed.per.capita>casespercapita.origin,1,0)) %>%
     filter(number.confirmed>0) %>%
     ungroup()

onsetday<-onsetday %>% 
     group_by(Country.Region) %>%
     arrange(Date) %>%
     mutate("day.since.onset"= cumsum(day.of.onset)) %>%
     filter(number.confirmed>0) %>%
     ungroup()

onsetday<- merge(onsetday, population, by=c("Country.Region"))

####################################################################### End of onset data

df.today<-df %>%
       group_by(Country.Region) %>%
       summarise(total.death=max(number.death),
                 percent.mortality=format(100*max(number.death)/max(number.confirmed),digits=3),
                 total.recovered=max(number.recovered),
                 total.confirmed=max(number.confirmed),
                 death.per.capita=as.numeric(format(max(number.death/population),digits=3)),
                 recovered.per.capita=as.numeric(format(max(number.recovered/population),digits=3)),
                 confirmed.per.capita=as.numeric(format(max(number.confirmed/population),digits=3)),
                 cases.per.capita=format((max(number.confirmed)-max(number.recovered)-max(number.death))/unique(population),digits=3),
                 population= unique(population),
                 continent=unique(continent))

   
df.today.above.cutoff <- df.today %>% filter(total.confirmed>confirmed.cutoff)

continental<-df %>%
             group_by(continent,Date) %>%
             summarise(number.confirmed=max(number.confirmed),
                       number.death=max(number.death),
                       number.recovered=max(number.recovered),
                       population=max(population)
                       ) 


### WORLDOMETHER TABLE 


df.table<-worldometer %>% 
          filter(Country.Other!="Total:")
df.table$Country.Other<-str_trim(as.character(df.table$Country.Other))

df.table$Country.Other[df.table$Country.Other=="UK"]<-"United Kingdom"

df.table$percent.mortality<-NA
df.table<-df.table[,c("Country.Other","ActiveCases","TotalCases","TotalRecovered","TotalDeaths","percent.mortality","Tot.Cases.1M.pop")]

cols = c(2:7);    
df.table[,cols] = apply(df.table[,cols], 2, function(x) as.numeric(sub(",", "",as.character(x)),fixed=T));

df.table$TotalDeaths[is.na(df.table$TotalDeaths)]<-0
df.table$TotalCases[is.na(df.table$TotalCases)]<-0
df.table$TotalRecovered[is.na(df.table$TotalRecovered)]<-0
df.table$percent.mortality<-df.table$TotalDeaths/df.table$TotalCases

df.table$Tot.Cases.1M.pop<-round(1000*df.table$TotalCases/df.table$Tot.Cases.1M.pop)*1000


df.table$Tot.Cases.1M.pop<-prettyNum(df.table$Tot.Cases.1M.pop, big.mark=",", scientific=F)
df.table$TotalCases<-prettyNum(df.table$TotalCases, big.mark=",")
df.table$TotalDeaths<-prettyNum(df.table$TotalDeaths, big.mark=",")
df.table$TotalRecovered<-prettyNum(df.table$TotalRecovered, big.mark=",")
df.table$percent.mortality<-percent(as.numeric(df.table$percent.mortality), accuracy=0.1, suffix="%")
df.table$Tot.Cases.1M.pop<-prettyNum(df.table$Tot.Cases.1M.pop, big.mark=",")
df.table$ActiveCases<-prettyNum(df.table$ActiveCases, big.mark=",")

names(df.table)[1]<-"<i class=\"fas fa-globe-americas \"></i>"
names(df.table)[2]<-"<i class=\"fas fa-procedures  \" title=\"Currently active COVID19 cases\"> </i> Active "
names(df.table)[3]<-"<i class=\"fas fa-procedures \" title=\"Total confirmed COVID19 cases\" > </i> Total"
names(df.table)[4]<-"Recovered"
names(df.table)[6]<-"<i class=\"fas fa-book-dead \" title=\"Total COVID19 fatalities\"> </i> Deaths"
names(df.table)[6]<-"<i class=\"fas fa-book-dead \" title=\"Percent of identified COVID19 cases that prove fatal\">&nbsp</i><i class=\"fas fa-percent\"></i>"
names(df.table)[7]<-"<i class=\"fas fa-male\" title=\"people in country\">  </i> Population"

df.today.above.cutoff$total.recovered.per.case<-df.today.above.cutoff$total.recovered/df.today.above.cutoff$total.confirmed
df.today.above.cutoff$total.death.per.case<-df.today.above.cutoff$total.death/df.today.above.cutoff$total.confirmed


# Get a list of highload countrie

highload  <- subset(df.today.above.cutoff,population>minimum.pop) %>%
              arrange(-confirmed.per.capita) 

highload<- head(highload$Country.Region,topcount)


######################################################### End of tidy                                                            


################################# End of setup
```


```{r setup, include=FALSE, echo=FALSE}
# Technical stuff for webpage
# Create header.html
CON <- file("header.html")
writeLines('<!-- Global site tag (gtag.js) - Google Analytics -->\n <script async src="https://www.googletagmanager.com/gtag/js?id=XXXXXXX"></script> \n <script> window.dataLayer = window.dataLayer || [];function gtag(){dataLayer.push(arguments);}gtag(\'js\', new Date()); gtag(\'config\', \'UA-157579559-2\');</script> <script src="https://kit.fontawesome.com/5961dcc1a3.js" crossorigin="anonymous"></script> <link href="https://fonts.googleapis.com/css?family=Exo+2:300&display=swap" rel="stylesheet"> <style> body {font-family: \'Exo 2\', sans-serif;font-size: 18px;}</style>', CON)
close(CON)
```
<br>
<div style="text-align: right"> <a href="nCOV19.html"><i class="fas fa-sync-alt" title="Refresh"></i></a> <a href="https://spendiff.com/"><i class="fas fa-home" color="lightgrey"></i>
</a> </div>
***

# Global COVID19 Data 
Last Updated: `r format(Sys.time(), format='%H:%M, %B %d, %Y (%Z)')` 

***

# Cases: <span style="color:steelblue">`r prettyNum(total.cases.worldometer, big.mark=",", scientific=F)`</span>

# Deaths: <span style="color:red">`r prettyNum(total.deaths.worldometer, big.mark=",", scientific=F)`</span>

# Recovered: <span style="color:darkgreen"> `r prettyNum(total.recovered.worldometer, big.mark=",", scientific=F)`</span>

## <span style="color:steelblue">`r format(100*growth$death[growth$Date==maxdate]/growth$confirmed[growth$Date==maxdate],digits=2)`%</span> of all (confirmed) cases end in death [^1]
</center>
<br>

<center>
 1 in `r round(1/(growth$death[growth$Date==maxdate]/growth$confirmed[growth$Date==maxdate]))` cases: `r toperson(round(1/(growth$death[growth$Date==maxdate]/growth$confirmed[growth$Date==maxdate])),"steelblue")` 
</center>
<br>
New Yesterday - Cases: <span style="color:steelblue">`r prettyNum(growth$newconfirmed[growth$Date==maxdate], big.mark=",", scientific=F)`</span>, Deaths: <span style="color:red">`r prettyNum(growth$newdeath[growth$Date==maxdate], big.mark=",", scientific=F)`</span>, Recovered: <span style="color:darkgreen"> `r prettyNum(growth$newrecovered[growth$Date==maxdate], big.mark=",", scientific=F)`</span> 

***

[^1]: For multiple reasons, the number of deaths divided by the number of cases is not the mortality rate of COVID19. It is the percentage of individuals identified to have COVID19 who die (or are dead). 


# Hot Spots


The 10 countries with the greatest number of **active cases** as of today (`r format(Sys.Date(), format='%B %d, %Y')`) are shown in the table below.
Data for all countries with active and historic cases are also accessible using the search box above the table.

<br><br>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=14,fig.height=8 }

datatable(df.table ,
          options = list( order = list(list(1,'desc')),                                  
                          columnDefs = list(list(className = 'dt-right', targets = 1:6) 
                                            )
                        ),
          rownames = FALSE, 
          escape=F,
          autoHideNavigation =T)
```
<br>

# Time Series Data

The following plots use a slightly slower time series dataset that may lag behind the live headline figures which are taken from a high-frequency data source.

## Propagation of Cases Geographically
 
Log-log Scatterplot (Räppli plot) of death per capita (aka per person) vs confirmed cases per capita for all countries currently with more than `r confirmed.cutoff` confirmed cases on `r format(Sys.Date(), format='%B %d, %Y')`. Colour and shape of the points indicate the continent. Size of the points gives a measure of the country population. Dotted lines indicate the boundaries for the percentages of cases that prove fatal (0.1%,1%,10% and 100%). 


              
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=7 }
days.data<-max(df$days.of.data)
# 
# #Make sure that all countries in the last frame are also in the first frame
# 

df.firstframe<- df %>% 
               group_by(continent) %>%
               arrange(Date) %>% 
               filter(row_number()==n())

df.firstframe$Country.Region<-""
df.firstframe$days.of.data<-0
df.firstframe$number.confirmed<-0
df.firstframe$population<-0

df.scatter<-subset(df, (number.death>0) & (days.of.data>0))

df.scatter<-bind_rows(df.firstframe,df.scatter)

gg <- ggplot(df.scatter) +
  geom_abline(aes(slope=1, intercept=0),colour="lightgrey",alpha=.8,linetype = "dashed") + 
  geom_abline(aes(slope=1, intercept=1),colour="lightgrey",alpha=.8,linetype = "dashed") + 
  geom_abline(aes(slope=1, intercept=2),colour="lightgrey",alpha=.8,linetype = "dashed") + 
  geom_abline(aes(slope=1, intercept=3),colour="lightgrey",alpha=.8,linetype = "dashed") + 
  geom_point(aes(number.death/population, 
                 number.confirmed/population,
                 color = continent,
                 size = population, 
                 shape = continent,
                 frame = days.of.data+1,
                 text=sprintf("Country: %s<br>Date: %s<br>Cases: %s <br>Deaths: %s<br>Death/Case: %s percent", Country.Region ,Date , number.confirmed, number.death,format(100*number.death/number.confirmed, digits=2)),
                 ids = Country.Region),
             alpha=.7) +
  scale_x_log10("Deaths per capita")  +
  scale_y_log10("Cases per capita")  +
  scale_size("")+
  scale_color_discrete("")+    
  scale_shape("")+  
  ggtitle("How we got to this: Confirmed cases and deaths per capita") +
  theme_light()
gg<-gg + annotate("text", x = .5e-8, y = .7e-5,colour="grey",label = "0.1%")
gg<-gg + annotate("text", x = .5e-8, y = .7e-6,colour="grey", label = "1%")
gg<-gg + annotate("text", x = .5e-8, y = .7e-7,colour="grey", label = "10%")
gg<-gg + annotate("text", x = .5e-8, y = .7e-8,colour="grey", label = "100%")

fig<- ggplotly(gg,tooltip="text")
fig<- fig %>% config(displayModeBar = FALSE)



fig <- fig %>%
  animation_slider(
    currentvalue = list(prefix = "Days of Data ", font = list(color="red"))
  )

fig


```

<br>

***

## Comparison of Spread

Take all of the countries with a population over `r prettyNum(casespercapita.cutoff, big.mark=",", scientific=F)` and look at cases relative to the first day where **confirmed** cases appear - 'onset' day. We examine how easily it continues to spread. Variations in detection efficiency mean that comparison of performance comes with caveats. 

To make plots more legible for specific continents, click  <i class="fas fa-mouse-pointer" style="color:steelblue"> </i> on individual continents in the plot legend to toggle their selection.


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9.8,fig.height=4 }

gg<-ggplot(onsetday)+
  geom_line(aes(day.since.onset,
                number.confirmed.per.capita,group=Country.Region,
                linetype=continent.y,
                colour=continent.y,
                text=sprintf("Country: %s<br>Date: %s<br>Cases: %s <br>Deaths: %s<br>Death/Case: %s percent", Country.Region ,Date , number.confirmed, number.death,format(100*number.death/number.confirmed, digits=2))))+
  scale_colour_discrete("")+
  scale_linetype_discrete("")+
  scale_y_continuous("cases per capita")+
  scale_x_continuous("Days since \'onset\'")+
  theme_light()

fig<-ggplotly(gg,tooltip="text")
fig<- fig %>% config(displayModeBar = FALSE)
#fig <- fig %>% layout(showlegend = FALSE)
fig

```

Disregarding the China data and looking at the increase in confirmed cases in all other countries.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9.8,fig.height=4 }

gg<-ggplot(subset(onsetday, Country.Region!="China"))+
  geom_line(aes(day.since.onset,
                number.confirmed - min(number.confirmed),group=Country.Region,
                linetype=continent.y,
                colour=continent.y,
                text=sprintf("Country: %s<br>Date: %s<br>Cases: %s <br>Deaths: %s<br>Death/Case: %s percent", Country.Region ,Date , number.confirmed, number.death,format(100*number.death/number.confirmed, digits=2))))+
  scale_colour_discrete("")+
  scale_linetype_discrete("")+
  scale_y_continuous("cases")+
  scale_x_continuous("Days since \'onset\'")+
  theme_light()

fig<-ggplotly(gg,tooltip="text")
fig<- fig %>% config(displayModeBar = FALSE)
#fig <- fig %>% layout(showlegend = FALSE)
fig

```

Disregarding the China data and looking at the COVID19 deaths in all other countries.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9.8,fig.height=4 }

gg<-ggplot(subset(onsetday, Country.Region!="China"))+
  geom_line(aes(day.since.onset,
                number.death - min(number.death),group=Country.Region,
                linetype=continent.y,
                colour=continent.y,
                text=sprintf("Country: %s<br>Date: %s<br>Cases: %s <br>Deaths: %s<br>Death/Case: %s percent", Country.Region ,Date , number.confirmed, number.death,format(100*number.death/number.confirmed, digits=2))))+
  scale_colour_discrete("")+
  scale_linetype_discrete("")+
  scale_y_continuous("death")+
  scale_x_continuous("Days since \'onset\'")+
  theme_light()

fig<-ggplotly(gg,tooltip="text")
fig<- fig %>% config(displayModeBar = FALSE)
#fig <- fig %>% layout(showlegend = FALSE)
fig

```


<br>

## Global Cases of COVID19

If we define Active cases as *Confirmed cases - Confirmed Recovery - Confirmed Death*, we get the plot below.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=4 }

growth.china.only<-growth.by.country %>% filter(Country.Region=="China")

growth$active<-growth$confirmed-growth$death-growth$recovered

fig <- plot_ly(type = 'scatter', mode = 'lines')


fig <- fig %>% add_trace(x = ~growth$Date, 
               y = ~growth$active, 
               hoverinfo = "text",
               fill = 'tozeroy',
               name = "All Countries",
               text = paste("Global<br>cases:",growth$confirmed,"<br>Deaths:",growth$death,"<br>Recovered:",growth$recovered,sep="")
)


fig <- fig %>% add_trace(
               x=~growth.china.only$Date,
               y= ~growth.china.only$active,
               fill = 'tozeroy',
               name = "China only",
               hoverinfo = "text",
               text = paste("China only<br>cases:",growth.china.only$confirmed,"<br>Deaths:",growth.china.only$death,"<br>Recovered:",growth.china.only$recovered,sep="")
               )

fig<- fig %>% config(displayModeBar = FALSE)
fig <- layout(fig, xaxis = list(title = "Date"),
                    yaxis = list(title = "Active cases"))

fig
  
```



The number of deaths per day


<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5,fig.height=4 }


fig <- plot_ly(hoverinfo = "text", text = paste("Date:",growth$Date,"<br>Deaths:",growth$newdeath,sep=""),type = 'scatter', mode = 'lines')
fig <- fig %>% add_trace( x = ~growth$Date,y = ~growth$newdeath, fill = 'tozeroy')
fig <- fig %>% layout(
    title = "deaths (daily)",
    xaxis = list(
      title = "Date",
      rangeselector = list(
        buttons = list(
           list(
            count = 7,
            label = "1 wk",
            step = "day",
            stepmode = "backward"),
           list(count = 1,
            label = "1 mo",
            step = "month",
            stepmode = "backward"),
           list(step = "all"))),

      rangeslider = list(type = "date",title = "Date")),

   yaxis = list(title = "Deaths"))
fig<- fig %>% config(displayModeBar = FALSE)
fig <- fig %>% layout(showlegend = FALSE)
fig
```

<br>

***
 
# Countries of (Personal) Interest

 `r coi`. 
 
 To make plots more legible for specific countries, click  <i class="fas fa-mouse-pointer" style="color:steelblue"> </i> on individual countries in the plot legend to toggle their selection.
 
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=6 }

timeplots<-subset(df, (Country.Region %in% coi))

fig<-ggplot(timeplots)+
  geom_line(aes(Date,(number.confirmed-number.recovered-number.death)/population, colour=Country.Region))+
  scale_colour_discrete("")+
  scale_y_continuous("Active cases per capita")+
  theme_minimal()

fig<-ggplotly(fig)
fig<- fig %>% config(displayModeBar = FALSE)
fig

```

New cases per day


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5,fig.height=4 }

timeplots<-subset(growth.by.country, (Country.Region %in% coi))

fig<-ggplot(timeplots)+
  geom_area(aes(Date,newconfirmed, fill=Country.Region), alpha=.5)+
  facet_wrap(~Country.Region, scales = "free_y", ncol=3)+
  scale_colour_discrete("")+
  scale_fill_discrete("")+
  scale_y_continuous("New confirmed cases")+
  theme_minimal()

fig<-ggplotly(fig)
fig<- fig %>% config(displayModeBar = FALSE)
fig <- fig %>% layout(showlegend = FALSE)
fig
```

<br>


***

## By Continent

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5,fig.height=4 }

gg<-ggplot(continental)+
    geom_area(aes(Date,number.confirmed, fill=continent),size=1, show.legend = FALSE, alpha=.7)+
    scale_y_continuous("cases")+
    ggtitle("Confirmed cases")+
    scale_fill_discrete()+
    facet_grid(.~continent)+
    theme_light()

gg + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The deaths-per-case may differ due to variations in detection efficiency. Poorer detection leads to a higher percentage.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5,fig.height=4 }

gg<-ggplot(continental)+
    geom_line(aes(Date,100*number.death/number.confirmed, colour=continent ),size=1, show.legend = FALSE, alpha=.7)+
    scale_y_continuous("Deaths per case (percent)")+
    ggtitle("Deaths per Case (%)")+
    scale_color_discrete()+
    facet_grid(.~continent)+
   theme_light()

gg + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

<br>

***

## Death and Recovery

The number or recoveries per case and death per case by country. For both of these metrics, the countries where cases occured earlier are likely to show higher numbers. 

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.3,fig.height=8 }


gg<- ggplot(df.today.above.cutoff) + 
         geom_bar(aes(x = reorder(Country.Region, df.today.above.cutoff$total.recovered.per.case), y = 100*df.today.above.cutoff$total.recovered.per.case, colour=continent, fill=continent),stat = "identity")+
         scale_y_continuous("Recovered per Case (%)")+
         scale_x_discrete("")+
         coord_flip()+  
        theme_minimal()
gg


```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.3,fig.height=8 }


gg<-ggplot(df.today.above.cutoff) + 
         geom_bar(aes(x = reorder(Country.Region, df.today.above.cutoff$total.death.per.case), y = 100*df.today.above.cutoff$total.death.per.case, colour=continent, fill=continent),stat = "identity")+
         coord_flip()+
         scale_y_continuous("Deaths per Case (%)")+
         scale_x_discrete("")+
         theme_minimal()

gg
```

cases in countries that are yet to register a death.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.3,fig.height=6 }


gg<-ggplot(subset(df.today.above.cutoff, total.death==0)) + 
            geom_bar(aes(x = reorder(Country.Region, total.confirmed), y = total.confirmed, colour=continent, fill=continent),stat = "identity")+
            coord_flip()+
            scale_y_continuous("cases")+
            scale_x_discrete("")+
            theme_minimal()

gg


```
<br>

## Deaths per Case 

Histogram of Death/Case for all countries where coronavirus deaths have occured and a population over `r prettyNum(minimum.pop, big.mark=",", scientific=F)`. The individual countries and death rate can be found using the hover-over on the last frame of [this plot](#propagation-of-cases-geographically).

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.3,fig.height=6 }


gg<-ggplot(subset(df.today.above.cutoff, total.death>0)) + 
            geom_histogram(aes(as.numeric(percent.mortality), fill=continent ),stat = "bin",bins = 10, show.legend = FALSE)+
            scale_y_continuous("countries")+
            scale_x_continuous("Cases leaing to death (%)")+
            facet_grid(continent~.)+
            theme_light()

gg


```
<br>

***

<br>


## Data Sources: 
- COVID19 Historical: `r urlcases`
- COVID19 Live: [Worldometers](https://www.worldometers.info/coronavirus/)
- population: `r url`
- geography: https://www.kaggle.com/statchaitya/country-to-continent